---
title: "Outbreak data from Sykdomspulsen"
output:
  pdf_document:
    toc: FALSE
    keep_tex: TRUE
header-includes:
  - \usepackage{booktabs}
  - \usepackage{longtable}
  - \usepackage{array}
  - \usepackage{multirow}
  - \usepackage{wrapfig}
  - \usepackage{float}
  - \usepackage{colortbl}
  - \usepackage{pdflscape}
  - \usepackage{tabu}
  - \usepackage{threeparttable}
  - \usepackage{threeparttablex}
  - \usepackage[normalem]{ulem}
  - \usepackage{makecell}
  - \usepackage{xcolor}
  - \makeatletter\renewcommand*{\fps@figure}{H}\makeatother
params:
  dev: TRUE
  package_dir: "."
editor_options: 
  chunk_output_type: console
---

```{r message=FALSE, warning=FALSE, include=FALSE, paged.print=FALSE}
fhi::DashboardInitialiseOpinionated("sykdomspuls",
                                    PACKAGE_DIR=params$package_dir,
                                    FORCE_DEV_PACKAGE_LOAD = params$dev,
                                    SILENT=TRUE)

suppressMessages(library(data.table))
suppressMessages(library(dplyr))
suppressMessages(library(ggplot2))
suppressMessages(library(kableExtra))
suppressMessages(library(fhiplot))
suppressMessages(library(lubridate))
suppressMessages(library(fhidata))
suppressMessages(library(gridExtra))
suppressMessages(library(grid))


byMunicipality <- TRUE
start_date = ymd("2019-05-29")
final_data_date =ymd("2019-06-13")
lineList <- fread(fhi::DashboardFolder("data_raw", "emerg_linelist_mult.txt"))
lineList[, KonsultasjonDato := ymd(KonsultasjonDato)]
population <- data.table(fhidata::norway_population_current[year == 2019, .(pop = sum(pop)), by=.(location_code)])
lineList[, kons_location_code := paste("municip", KonsKommuneNr, sep="")]
lineList[, pat_location_code := paste("municip", PasientKommuneNr, sep="")]

dates = unique(lineList[, KonsultasjonDato])

all_dates = seq(min(dates), max(dates), "days")

frac_consultations <- sin(0:(length(all_dates) - 1) / length(all_dates) * pi)


subchunkify <- function(g, fig_height=7, fig_width=5) {
  g_deparsed <- paste0(deparse(
    function() {g}
  ), collapse = '')
  
  sub_chunk <- paste0("
  `","``{r sub_chunk_", floor(runif(1) * 10000), ", fig.height=", fig_height, ", fig.width=", fig_width, ", echo=FALSE, fig.fullwidth=TRUE}",
  "\n(", 
    g_deparsed
    , ")()",
  "\n`","``
  ")
  
  cat(knitr::knit(text = knitr::knit_expand(text = sub_chunk), quiet = TRUE))
}


```
\begin{center}
```{r echo=FALSE, results='asis', paged.print=TRUE }
cat(paste("Data updated on the", format(final_data_date)))
```
\end{center}
# Outbreak Overview

```{r echo=FALSE, results='asis', paged.print=TRUE }

 nCases = nrow(lineList[KonsultasjonDato > ymd("2019-05-29")])
 cat("Data from the syndromic surveillance system Sykdomspulsen related to outbreak in Askøy and surrounding areas (please note that thsi draft is not using real data). From the 29th of May we have registered ", nCases, " consultations potentially related to the outbreak. This report includes data until the 13th of June \n")


```

\medskip

# Time evolution of the outbreak

The chart shows the number of consultations with gastro-intestinal or similar symptoms per day as received by Sykdomspulsen.

```{r echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth=TRUE, results='asis'}
q <- ggplot(lineList[, .(N = .N), by=.(KonsultasjonDato)], aes(x=KonsultasjonDato, y=N)) +
     geom_col(fill = fhiplot::base_color, width = 0.8) +
     theme_fhi_lines() +
     ggtitle("Outbreak consultations per day")+
     ylab("")
subchunkify(q, 5, 8)
```





```{r echo=FALSE, message=FALSE, warning=FALSE, fig.fullwidth=TRUE, results='asis'}
## cat("The chart shows the fraction of consultations with gastro-intestinal or similar symptoms from Sykdomspulsen.")

## q1 <- ggplot(NULL, aes(x=all_dates, y=frac_consultations)) +
##      geom_line(color = fhiplot::base_color, size=2) +
##      scale_y_continuous(labels=scales::percent) +
##      ylab("Percent of consultations") +
##      ggtitle("Fraction of reported gastro-intestinal consultations") +
##      theme_fhi_lines() +
##      xlab("")
## subchunkify(q1, 5, 8)


```




# Geographic distribution


## Municipality of the consultation





```{r echo=FALSE, results='asis', paged.print=TRUE, fig.fullwidth=TRUE, out.width=c('50%', '50%'), fig.show='hold' }

create_map_table <- function(data, column_name, column_code, out_col="N", irFactor = 10000){
  # Creates a table and a map in  a two-column layout
  if(out_col == "N"){
    plot_title = "Consultations by Municipality"
    table_col_title = "Consultations"
    plot_color_title = "Number of Consultations"
  }else if(out_col == "ir"){
    table_col_title = paste("Incidence Rate (per", irFactor, ")") 
    plot_title = "Incidence Rate by Municipality"
    plot_color_title = "Incidence Rate"
 }
  municip_cases <- data[, .(N=.N), by=.("name"=get(column_name), "location_code"=get(column_code))]  

  municip_cases <- population[municip_cases, on=.(location_code =location_code)]
  municip_cases[, ir:= round(N  /pop * irFactor)]

  table <- municip_cases[order(-get(out_col)), .(name, get(out_col))]

  cat(knitr::raw_latex("\\begin{minipage}{0.39\\textwidth}"))
  setnames(table,c(
    "Municipality", table_col_title)
  )
  k <- knitr::kable(table, "latex",
                      booktabs = T,
                      align = "c",
                      linesep = "")
   print(k)
  cat(knitr::raw_latex("\\end{minipage}
  \\begin{minipage}{0.69\\textwidth}"))


 max <- max(municip_cases[, get(out_col)])
 municip_cases[, binned := cut(get(out_col), round(max/5 * 0:5))]

 pd <- fhidata::norway_map_municips

 plot_data <- data.table(pd)[municip_cases, on=.(location_code=location_code), nomatch=0]

 cnames <- aggregate(cbind(long, lat) ~ location_code, data=plot_data, FUN=function(x)mean(range(x)))
 cnames <- inner_join(cnames, fhidata::norway_locations_current, by=c("location_code" = "municip_code" ))

 q <- ggplot() +
   geom_polygon(data = plot_data, aes( x = long, y = lat, group = group, fill=binned),
                color="black") +
   theme_void() +
   coord_quickmap() +
   ggtitle(plot_title) + 
   fhiplot::scale_fill_fhi(plot_color_title, palette = "map_seq_complete", direction = 1) +
 geom_label(data=cnames, aes(long, lat, label = municip_name), size=3)

 subchunkify(q, 5, 5)
 cat(knitr::raw_latex("\\end{minipage}"))
}
create_map_table(lineList, "KonsKommune", "kons_location_code")
create_map_table(lineList, "KonsKommune", "kons_location_code", out_col="ir")

```



## Municipality where the patient lives

```{r echo=FALSE, results='asis', paged.print=TRUE }
create_map_table(lineList, "PasientKommune", "pat_location_code")
```

\newpage

# Demographics and details

We show a breakdown of the outbreak by demographics, diagnosis and type of consultation. 


```{r echo=FALSE, message=FALSE, warning=FALSE, results='asis', fig.fullwidth=TRUE}


table_chart <- function(data, column){
  cat("\\subsection{", column, "}\n")
  cat("Breakdown of all the cases from ", format(start_date), "\n")
  breakdown <- data[, .(N=.N), by=.(get(column))]
  table <- breakdown[order(-N), .(get, N)]
  setnames(table,c(
    column,"Cases")

  )

  k <- knitr::kable(t(table), "latex",
                      booktabs = T,
                      align = "c",
                      linesep = "")
   print(k)

  cat("Timeline of cases by", column, "\n")
  data2 <- data[, .(N = .N), by=.(KonsultasjonDato, get(column))]
  # Hack as data.table returns get as column name
  if (length(unique(data2[, get])) > 5){
    q <- ggplot(data2, aes(x=KonsultasjonDato, y=N)) +
        lemon::facet_rep_wrap(~get, repeat.tick.labels = T) +
        geom_col(fill = fhiplot::base_color, width = 0.8) +
        fhiplot::scale_fill_fhi(column ,palette = "primary", direction = 1) + 
        theme_fhi_lines() +
        ylab("") +
        theme(axis.text.x = element_text(angle = 30))

   subchunkify(q, 9, 9)
 
  } else {
    q <- ggplot(data2, aes(x=KonsultasjonDato, y=N, fill=get))+
        geom_col(width = 0.8) + 
        fhiplot::scale_fill_fhi(column ,palette = "primary", direction = 1) + 
        theme_fhi_lines() +
        ylab("")
   subchunkify(q, 4, 8)
  }
  cat(" \\newpage ")
}


columns <- c("Kjønn", "Alder", "DiagnoseBeskrivelse", "Kontaktype", "Praksis")

if(byMunicipality){
  newColumns <- c("KonsKommune", "PasientKommune", columns)
  for(column in newColumns){
     table_chart(lineList, column)
  }

 municip_cases <- lineList[, .(N=.N), by=.(KonsKommune, kons_location_code)]
 table <- municip_cases[order(-N), .(KonsKommune, N)]
 for(i in 1:nrow(table)){
  municipName = table[i, KonsKommune]
  cat("\\section{", municipName, "}\n")

  ## fc <- frac_consultations*runif(0.8,1, length(frac_consultations))
  ## q1 <- ggplot(NULL, aes(x=all_dates, y= frac_consultations)) + 
  ##    geom_line(color = fhiplot::base_color, size=2) +
  ##    scale_y_continuous(labels=scales::percent) +
  ##    ylab("Percent of consultations") +
  ##    ggtitle("Fraction of reported gastro-intestinal consultations")+
  ##    theme_fhi_lines() + xlab("")
  ## subchunkify(q1, 5, 8)


  for(column in columns){
     table_chart(lineList[KonsKommune == municipName], column)
    }


  }



} else {
  for(column in columns){
     table_chart(lineList, column)
  }

}

```
