---
title: "Input data"
author: "Richard White"
date: "2018-10-12"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEncoding{UTF-8}
  %\VignetteEngine{knitr::rmarkdown}
editor_options: 
  chunk_output_type: console
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
library(sykdomspuls)
```

`sykdomspuls` requres a specific raw data format. We require one (or less) rows for each structural variable:

```{r}
print(as.character(sykdomspuls::VARS$REQ_DATA_RAW_STRUCTURAL))
```

For each row, the variables saved in `sykdomspuls::CONFIG$SYNDROMES` and `sykdomspuls::VARS$REQ_DATA_RAW_OTHER` contain the number of consultations that occurred for that combination of structural variables.

```{r}
print(as.character(sykdomspuls::CONFIG$SYNDROMES))
print(as.character(sykdomspuls::VARS$REQ_DATA_RAW_OTHER))
```

In total, we require the following variables: 

```{r}
print(as.character(sykdomspuls::VARS$REQ_DATA_RAW))
```

We can use the function `GenFakeDataRaw()` to generate a set of fake data for `municip5054`.

```{r}
d <- GenFakeDataRaw()
print(d)
```

